---
title: "Baseline Model Exploration"
date: "3/26/2021"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(ggcorrplot)
library(GGally)

options(scipen = 999)

#read in 14 days .csv file
working_df <- read.csv("working_14days.csv")
```



##Baseline Model Analysis

## Model 1: Regularized (L1 - LASSO) Linear Regression Model



### Step 1: Clean up the data on the demographic indicators

### Check for Collinearity

Step 2: Select 14 days before and 14 days after the first policy was implemented for the time range
Step 3: Split the data into a training and testing set
Step 4: Apply LASSO regularization (L1 regularization) to select the important control variables 
Step 5: Re-run linear regression to get coefficient estimates on the variables that are interpretable
Step 6: After the control-only baseline model was run, then the user would be able to select the non-pharmaceutical interventions that they want 




```{r}

names(working_df)

#do we need the date of the closures/openings once we've used them to create the binary variables?
#there is some demographic information that I think that we can eliminate as well


#how do we take the time dimension into account?

# working_df_subset %>% 
  working_df

```








